---
title: "Mini-Project 2"
author: "Team Hedgehog!"
date: "2/8/2021"
output:
  html_document:
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)
library(tidyverse)
library(mosaic)
```

```{r}

## M -- taking in raw data, this is updated as of feb 9th at 7:30 am ish

raw_data <- read_csv("./raw_data.csv")
```

```{r}

## M - Cleaning data part

cleaned_data <- raw_data %>%
  filter(Consent == "Yes, I consent" & Age == "Yes, I am 18 or older") %>%
  select(Happiness1_1, `Animal-Preference`, Happiness2_1, animal) %>%
  rename(preferred_animal = `Animal-Preference`,
         trial_1 = Happiness1_1,
         trial_2 = Happiness2_1) %>%
  mutate(subject_id = as.factor(row_number())) %>% 
  na.omit() %>% # K, response to M - here's a faster way to remove incomplete cases. We definitely removed that 1 person, and just that 1 person! :)
  
  ## M - BIG NOTE that there was one missing value for the second happiness question, and I removed it here (it was a roundabout way of doing it, I know, but I don't think we'll be graded on that particular point)
  
  ## MAKE sure to report that we deleted the person
  
# syntax from https://tidyr.tidyverse.org/reference/pivot_longer.html
  
  pivot_longer(cols = c(trial_1, trial_2), names_to = "trial", values_to = "happiness") %>%
  mutate(happiness = as.numeric(happiness))
```

```{r, eval = FALSE}

## M - saving clean data for submission purposes

write.csv(x = cleaned_data, file = "cleaned_mp2.csv", sep = ",")
```


## Working with within-subjects var -- trial would be the within-subjects variable here


```{r}

## M - SDs in animal/preferred animal groups; SDs in trial/animal/preferred animal groups do in fact meet the same standard deviation assumption

cleaned_data %>%
  group_by(animal, preferred_animal) %>%
  summarize(
    mean = mean(happiness),
    sd = sd(happiness)
  )

## check within too just to be sure

cleaned_data %>%
  group_by(trial) %>%
  summarize(
    mean = mean(happiness),
    sd = sd(happiness)
  )
```

same sd assumption met

```{r}
## M - basic boxplot with within-subjects var, with data that does not violate assumptions

ggplot(cleaned_data, aes(x = trial, y = happiness, color = preferred_animal)) +
  geom_boxplot()

ggplot(cleaned_data, aes(x = trial, y = happiness, color = animal)) +
  geom_boxplot()
```

```{r}
## Sarah - Basic Boxplot for Trial

ggplot(cleaned_data, aes(x = trial, 
                         y = happiness)) +
  geom_boxplot(fill = "seagreen3",
               color = "grey30") +
  labs(title = "Cute Animals and Happiness", 
       x = "Measurement Timeslot", 
       y = "Level of Happiness") +
  scale_x_discrete(labels = c('Before',
                              'After')) +
  scale_y_continuous(labels = c('0',
                                '2.5',
                                '5',
                                '7.5',
                                '10'))
```


```{r}
## Within-subjects scatterplot
for_scatterplot <- cleaned_data %>%
  pivot_wider(names_from = trial, values_from = happiness)

ggplot(data = for_scatterplot, aes(x = trial_1, y = trial_2)) +
    geom_point()
```


constant effects, additive assumptions met
maybe ANOVA

```{r}
# ANOVA
mod <- aov(happiness ~ preferred_animal*animal*trial + Error(subject_id) + preferred_animal*animal, data = cleaned_data)

summary(mod)

# yay, we made people happy!
```


```{r}
cleaned_data <- cleaned_data %>%
  mutate(fitted = fitted(mod), 
         residuals = residuals(mod))

ggplot(cleaned_data, aes(x = fitted, residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red")
```


## effect sizes

Overall variance explained by factor of interest: eta^2

$$\eta^2 = \frac{MS_{treatment}}{MS_{total}}$$
```{r}
# trial
# etasq <- 6.638 / 
# what is MSTotal?
```


## confidence intervals

alpha = 0.05 level

```{r}
# confidence interval
# compare trials (within-blocks)
# happiness 1 vs happiness 2
stats <- favstats(happiness ~ trial, data = cleaned_data)[,c(1,7,9)]
# diff of means
diff <- stats[1,2] - stats[2,2]
# use counts in leverage
n1 <- stats[1, 3]
n2 <- stats[2, 3]
leverage <- sqrt((1/n1)+1/n2)
# within-blocks error is MS_Res - fill by hand for simplicity
MSError <- 0.268
SE <- leverage * MSError

# df from error term
df <- 122
t_star <- qt(0.95, df, lower.tail = TRUE)

# confidence interval
lower <- diff - SE * t_star
upper <- diff + SE * t_star

# we are 95% confident that the difference between trial 1 and trial 2 is between lower and upper (-0.381, -0.269)
```


```{r}
# confidence interval
# compare cats and hedgehogs for trial 2
stats <- favstats(happiness ~ trial + animal, data = cleaned_data)[,c(1,7,9)]
# diff of means
diff <- stats[2,2] - stats[4,2]
# use counts in leverage
n1 <- stats[2, 3]
n2 <- stats[4, 3]
leverage <- sqrt((1/n1)+1/n2)

# MSPooled
# sum of error SSs / sum of error dfs
MSError <- (1049.3 + 32.64) / (122 + 122)
SE <- leverage * MSError

# df is (sum of error SSs)^2/sum of (sS^2/df)
df <- (1049.3 + 32.64)^2 / (((1049.3^2)/122) + ((32.64^2)/122))
t_star <- qt(0.95, df, lower.tail = TRUE)

# confidence interval
lower <- diff - SE * t_star
upper <- diff + SE * t_star

# we are 95% confident that the difference between animals for trial 2 is between lower and upper (-1.053, 1.564)
# this contains 0!!!
```


# Code Archive

## Working with happiness change, no within-subjects var


```{r}

## M - Cleaning data part

cleaned_data <- raw_data %>%
  select(Consent, Age, Happiness1_1, `Animal-Preference`, Happiness2_1, animal) %>%
  rename(preferred_animal = `Animal-Preference`,
         consent = Consent,
         age = Age) %>%
  filter(consent == "Yes, I consent" & age == "Yes, I am 18 or older") %>%
  mutate(happiness_change = as.numeric(Happiness2_1) - as.numeric(Happiness1_1)) %>%
  select(-consent, -age) %>%
  mutate(id = as.factor(row_number())) %>%
  na.omit() # K, response to M - here's a faster way to remove incomplete cases. We definitely removed that 1 person, and just that 1 person! :)
  
  ## M - BIG NOTE that there was one missing value for the second happiness question, and I removed it here (it was a roundabout way of doing it, I know, but I don't think we'll be graded on that particular point)
  
  ## MAKE sure to report that we deleted the person
```

```{r, eval = FALSE}

## M - saving clean data for submission purposes

write.csv(x = cleaned_data, file = "cleaned_mp2.csv", sep = ",")
```

```{r}

## M - putting the data in within subjects var shape w/ pivot

alt_data <- cleaned_data %>%
  select(-happiness_change) %>%
  rename(trial_1 = Happiness1_1,
         trial_2 = Happiness2_1) %>%
  
# syntax from https://tidyr.tidyverse.org/reference/pivot_longer.html
  
  pivot_longer(cols = c(trial_1, trial_2), names_to = "trial", values_to = "happiness") %>%
  mutate(happiness = as.numeric(happiness))

```

Michelle's notes, please read

So here's the story folks: I ran the "Working with happiness change, no within-subjects var" section and noticed that the same SD assumption was violated, like, barely, but enough to be a problem. So, I have two alternatives here, pre meeting with Randi. 

The first is Kathleen's transformation option, which is cool. It does not work to mitigate the S assumption violation. It also makes interpretation kind of seriously funky. 

So, the second option is to use trial as a within-subjects variable. It seems to fix the assumption violation, but of course it also complicates our lives significantly in other ways. 

```{r}

## M - finding the standard deviations of all groups for happiness_change version

cleaned_data %>%
  group_by(animal, preferred_animal) %>%
  summarize(
    mean = mean(happiness_change),
    sd = sd(happiness_change)
  )
```   

```{r}
## M - basic boxplot with happiness_change, no within-subjects var, with data that violates assumptions

ggplot(cleaned_data, aes(x = animal, y = happiness_change, color = preferred_animal)) +
  geom_boxplot()
```


## Kathleen's transformation suggestion to work with happiness_change

```{r, message = FALSE, warning = FALSE}
transformed_data <- cleaned_data %>%
  mutate(
    lhappiness_change = log(happiness_change + 10)
  )

transformed_data %>%
  group_by(animal, preferred_animal) %>%
  summarize(
    mean = mean(lhappiness_change),
    sd = sd(lhappiness_change)
  )
```



